---
title: Exploratory Data Analysis of P2P-Lending sites
subtitle: Analysing Loan Data From Prosper.com
author: Frederik Tischbein
date: November 3rd, 2018
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
---
```{r echo=FALSE, include=FALSE, Setup}
knitr::opts_chunk$set(echo = TRUE, message =FALSE, warning=FALSE, cache=FALSE)
```

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
library(dplyr)
library(tidyr)
library(reshape2)
library(ggplot2)
library(maps)
library(gridExtra)
library(scales)
library(lubridate)
library(rockchalk)
library(RColorBrewer)
library(GGally)

theme_set(theme_minimal())
```

```{r echo=FALSE, Load_the_Data}
prosper <- read.csv("prosperLoanData.csv") 
```

## Introduction

Propser is a peer-to-peer lending marketplace. Borrowers apply online for a 
fixed-rate, fixed-term loan between 2000 and $40,000 (here 1,000 and 35,000). 
Individuals and 
institutions can invest in the loans and earn attractive returns. Prosper 
handles all loan servicing on behalf of the matched borrowers and investors.

In this analysis I will focus on the question, which factors influence the 
borrowing rate an individual recieves when applying for a loan on Prosper.com.
My exploration and analysis of the different variables in the dataset and their
relationship are the basis for a linear model, I introduce in the second part, 
which can predict the borrowing rate an indivudal will recieve.

## Structure of the dataset
```{r echo=FALSE}
glimpse(prosper)
```

The dataset set contains 113,937 observations with 81 variables, including 
the loan amount, borrower rate (or interest rate), current loan status, borrower 
income, borrower employment status, borrower credit history, and the latest 
payment information. A detailed description of all the variables in the dataset 
can be found [here](https://rstudio-pubs-static.s3.amazonaws.com/86324_ab1e2e2fa210452f80a1c6a1476d7a2a.html).

## Data Wrangling & Cleaning
Looking at the structure of the dataset, we can notice a couple of things we 
need to fix in order to analyze the data more easily:

* According to the Prosper API documentation, the `ListingCategory..numeric.` 
variable describes the Listing category the borrower selected when posting their 
listing. We want to replace the numeric value with the category belonging to 
this number.
* Secondly, we need to adjust the structure of the `LoanOriginationDate` and the
`LoanOriginationQuarter` columns to be able to analyze the data over time. The 
`lubridate`, `dplyr` and `tidyr`packages come in handy here.
* And third, we combine two variables of creditworthyness, the `CreditGrade`
(used for loans pre 2009) and the `ProsperRating(alpha)`(used for loans
originated after July 2009) into one column called `CreditRating`.

```{r Cleaning Listing Categories}
# Replacing the numeric listing category with the category defined by the API
code2category <- function(code) {
  categories = c("Not Available", "Debt Consolidation", "Home Improvement",
                 "Business", "Personal Loan", "Student Use", "Auto", "Other",
                 "Baby&Adoption", "Boat", "Cosmetic Procedure", 
                 "Engagement Ring", "Green Loans", "Household Expenses",
                 "Large Purchases", "Medical/Dental", "Motorcycle", "RV",
                 "Taxes", "Vacation", "Wedding Loans")
  categories[2+1]
  return (categories[(code + 1)]) # in R Vector index starts at 1 not 0
}
prosper$ListingCategory <- sapply(prosper$ListingCategory, code2category)
```

```{r Cleaning Time data}
# Parsing the LoanOriginationDate column into the "date"-Format
prosper$LoanOriginationDate <- as_date(prosper$LoanOriginationDate)


# Transforming LoanOriginationQuarter from [Q# Year] to [Year Q#]
# This will make sorting and plotting much easier
prosper$LoanOriginationQuarter <- as.character(prosper$LoanOriginationQuarter)
prosper <- prosper %>%
            separate(col = LoanOriginationQuarter,
                      into = c("Quarter", "Year"), sep = " ") %>%
            unite(col = LoanOriginationQuarter, Year, Quarter, sep = " ")

# Making a LoanOriginationQuarterF column
prosper$LoanOriginationQuarterF <- factor(prosper$LoanOriginationQuarter)

# Due to the transformed LoanOriginationQuarter column the standard order is
# Q1 2006 Q2 2006 Q3 2006 Q4 2007 Q1 2007...
```

```{r Cleaning Credit Ratings}
# Joining the ProsperScore(alpha) and the CreditGrade into one column
prosper <- unite(prosper, col = "CreditRating", 
                 c("CreditGrade", "ProsperRating..Alpha."), sep = "")
prosper$CreditRating <- as.character(prosper$CreditRating)

# Setting CreditRating with the values "" or NC to NA
prosper$CreditRating[prosper$CreditRating == ""] <- NA
prosper$CreditRating[prosper$CreditRating == "NC"] <- NA

# Turning the variable into an ordered factor
prosper$CreditRating <- ordered(prosper$CreditRating, 
                              labels = c("AA","A","B","C","D", "E", "HR"))
```

```{r Cleaning_LoanStatus, include=FALSE}
# Collapsing all Loans that are 'Past due' into one category
loanStatusAdj <- combineLevels(prosper$LoanStatus,
  levs = c(
    "Past Due (>120 days)",
    "Past Due (16-30 days)",
    "Past Due (61-90 days)",
    "Past Due (1-15 days)",
    "Past Due (31-60 days)",
    "Past Due (91-120 days)"
  ),
  newLabel = c("Past Due")
)

loanStatusAdj <- as.character(loanStatusAdj)
loanStatusAdj[is.na(loanStatusAdj)] <- "Past due"

# Reordering the factor levels
loanStatusAdj <- factor(loanStatusAdj, levels = c("Cancelled", 
                                                  "Chargedoff", 
                                                  "Past due", 
                                                  "Defaulted", 
                                                  "Current", 
                                                  "FinalPaymentInProgress", 
                                                  "Completed"))
loanStatusAdj[is.na(loanStatusAdj)] <- "Past due"
```

## Univariate Plots
Great, let's explore the dataset! First, let's look at the size and status of 
loans issued on prosper.com:

Summary of Loan amounts:
```{r echo=FALSE, LoanAmount_Summary}
summary(prosper$LoanOriginalAmount)
```
Plots:
```{r echo=FALSE, Univariate_Plots1, fig.dim = c(10,4), fig.align="center"}
p1 <- ggplot(prosper, aes(x = LoanOriginalAmount)) +
      geom_histogram(binwidth = 1000) +
      ggtitle("Loans by Loan Amount") +
      scale_x_continuous(breaks = seq(0, 35000, 5000))

p2 <- ggplot(filter(as.data.frame(loanStatusAdj), loanStatusAdj!="Cancelled" & loanStatusAdj!="FinalPaymentInProgress"), aes(x = loanStatusAdj)) +
  geom_bar(aes(y = (..count..) / sum(..count..)),
    show.legend = FALSE
  ) +
  scale_y_continuous(
    breaks = seq(0, 0.5, 0.05),
    labels = percent
  ) +
  labs(
    y = "", x = "Loan Status",
    title = "Loans by Loan Status"
  )
grid.arrange(p1,p2, ncol=2)

```

* The issued loans range from $1000 to $35,000. However, we can see that the most 
popular loan sizes are $5000, §15,000 and $10,000.
* As we can see, 33% of loans in the dataset have been completed. About half 
(50%) of the issued loans in the dataset are currently running with all payments 
beeing made on time. 
* Loans are marked as "Defaulted" for reasons such as bankruptcy of the 
borrower. 
* About 2% of loans are marked as "Passed due", meaning that the last payment is
delayed between 1 and 120 days.
* After 4 missed payments /i.e. 120 days) a loan is marked as "Chargedoff". 
The likelyhood of recovering the debt is assumed to be very low. This affects 
about 10% of issued loans.

But for what kind of investements do people take a loan?

```{r echo=FALSE, Univariate_Plots2, fig.dim = c(10,5), fig.align="center"}
p1 <- ggplot(prosper, aes(x = reorder(ListingCategory,ListingCategory,
                     function(x)+length(x)))) +
  ggtitle("Loans By Listing Categories") +
  geom_bar() +
  coord_flip() +
  xlab("Loan Category")

# Adjusting the order of IncomeRange levels
prosper$IncomeRange <- factor(prosper$IncomeRange, 
                              levels = c("Not displayed",
                                         "Not employed",
                                         "$0",
                                         "$1-24,999",
                                         "$25,000-49,999",
                                         "$50,000-74,999",
                                         "$75,000-99,999",
                                         "$100,000+"))

p2 <- ggplot(prosper, aes(x = IncomeRange)) +
  ggtitle("Loans By Borrower's Income Range") +
  geom_bar() +
  coord_flip()

grid.arrange(p1,p2, ncol=2)
```

* Debt Consilidation (i.e. a form of debt refinancing that entails taking out 
one
loan to pay off many others), is by far the most popular reason for people to 
take on a loan. 
In fact, as explained in [this](https://www.forbes.com/sites/andrewjosuweit/2017/09/09/how-to-use-a-peer-to-peer-loan-to-pay-off-high-interest-debt/#5c2459551f79) article it is an attractive option to
pay-off high interest debt, such as credit cards.
* In the plot on the right, we can see that most of the borrowers on prosper 
have a lower-medium to medium income.
* Let's have a closer look at the monthly income of users:

```{r echo = FALSE, Monthly_Income}
# Plot remove outliers, log10 transformation
ggplot(prosper, aes(x = StatedMonthlyIncome)) +
  geom_histogram(binwidth = 500) +
  scale_x_continuous(lim = c(0, quantile(prosper$StatedMonthlyIncome, 0.99)),
                     breaks = seq(0, 21000, 2000))

summary(prosper$StatedMonthlyIncome)
```

* The median income of borrower is 4667. The distribution is right-skewed with 
outliers earning over a million a month.
* Lets look at monthly Loan payments:

```{r echo=FALSE, Monthly_Loan_Payment}
ggplot(prosper, aes(x = MonthlyLoanPayment)) +
  geom_histogram(binwidth = 50) +
  scale_x_continuous(lim = c(0, quantile(prosper$MonthlyLoanPayment, 0.99)),
                     breaks = seq(0, 875, 50))
```

* Most common is a monthly loan payment of 100-150$. Let's set the monthly
payment into relation to the monthly income, i.e. the percentage of the 
monthly income payed for the loan:

```{r echo=FALSE, Loan2Income}
p_loan_income <- prosper$MonthlyLoanPayment / prosper$StatedMonthlyIncome
summary(p_loan_income)
```

* Most loan payments range between 2.9 and 7.9 % of monthly
income, with the median beeing about 5%.
000,10000,15000,20000,30000))

## Plotting Time-Series
```{r echo=FALSE, fig.dim = c(12,5), fig.align= "center"}
loan_over_time <- prosper %>%
                  group_by(LoanOriginationQuarterF) %>%
                  summarise(sum_loans_originated = sum(LoanOriginalAmount))

# Dropping last Quarter because not fully recorded
loan_over_time <- subset(loan_over_time, LoanOriginationQuarterF != "2014 Q1")


ggplot(loan_over_time, aes(x = LoanOriginationQuarterF,
                           y = sum_loans_originated,
                           group = 1)) +
  geom_bar(stat = "identity") +
  ggtitle("Loan Amount Originated Per Quarter") +
  scale_y_continuous(
    labels = dollar_format(scale = 0.000001, suffix = "M")
  ) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

* Clearly, there is an upwards trend of loans beeing originated on prosper.
In 2013, growth has been very especially fast.
* There is a lowpoint in Q2 2009. A possible explanation for this is that 
Prosper relaunched in July 2009, after having obtained SEC registration for its 
loans.


```{r echo=FALSE, Defaulted_Loans}
defaulted_over_time <- subset(prosper, LoanStatus == "Defaulted") %>%
                  group_by(LoanOriginationQuarterF) %>%
                  summarise(n_defaulted = n())

ggplot(defaulted_over_time, aes(x = LoanOriginationQuarterF,
                           y = n_defaulted,
                           group = 1)) +
  geom_line(size = 1.1) +
  ggtitle("Number of Defaulted Loans per Quarter") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

* In the early years of Prosper, the number of defaulted loans was higher.

## Bivariate Plots

```{r Loan Issuance By State, echo=FALSE, fig.align="center"}
# Function to turn state abreviation into corresponding name
abbr2state <- function(abbr) {
  ab    <- tolower(c("AL",
             "AK", "AZ", "KS", "UT", "CO", "CT",
             "DE", "FL", "GA", "HI", "ID", "IL",
             "IN", "IA", "AR", "KY", "LA", "ME",
             "MD", "MA", "MI", "MN", "MS", "MO",
             "MT", "NE", "NV", "NH", "NJ", "NM",
             "NY", "NC", "ND", "OH", "OK", "OR",
             "PA", "RI", "SC", "SD", "TN", "TX",
             "CA", "VT", "VA", "WA", "WV", "WI",
             "WY", "DC"))
  st    <- c("Alabama",
             "Alaska", "Arizona", "Kansas",
             "Utah", "Colorado", "Connecticut",
             "Delaware", "Florida", "Georgia",
             "Hawaii", "Idaho", "Illinois",
             "Indiana", "Iowa", "Arkansas",
             "Kentucky", "Louisiana", "Maine",
             "Maryland", "Massachusetts", "Michigan",
             "Minnesota", "Mississippi", "Missouri",
             "Montana", "Nebraska", "Nevada",
             "New Hampshire", "New Jersey", "New Mexico",
             "New York", "North Carolina", "North Dakota",
             "Ohio", "Oklahoma", "Oregon",
             "Pennsylvania", "Rhode Island", "South Carolina",
             "South Dakota", "Tennessee", "Texas",
             "California", "Vermont", "Virginia",
             "Washington", "West Virginia", "Wisconsin",
             "Wyoming", "District of Columbia")
  tolower(st[match(tolower(abbr), ab)])
}

state_data <- prosper %>%
  group_by(BorrowerState) %>%
  summarise(total_loans_issued = sum(LoanOriginalAmount))

map <- map_data("state")
ggplot(state_data, aes(fill = total_loans_issued)) +
  ggtitle("Loan Issuance By State") +
  geom_map(aes(map_id = abbr2state(BorrowerState)), map = map) +
  expand_limits(x = map$long, y = map$lat) +
  scale_fill_distiller(
    direction = 1,
    guide = "legend",
    name = "Total Loans",
    labels = dollar_format(scale = 0.000001, suffix = "M")) +
  theme_void() +
  theme(plot.title = element_text(hjust = 0.5))
```

* As we can see, the state with the highest total of loans issued so far is 
California (> $125M).
* On second place there is Texas, New York and Florida which all have araound 55
-60M $ of total loans.

What type of rates are borrowers getting on prosper.com?

```{r echo=FALSE, AvgBorrowerRateBYCreditRating, fig.width=12}
rate_by_CreditRating <- subset(prosper, !(is.na(CreditRating))) %>%
  group_by(LoanOriginationQuarter, Term, CreditRating) %>%
  summarise(avg_rate = mean(BorrowerRate))

ggplot(rate_by_CreditRating, aes(x = LoanOriginationQuarter,
                                 y = avg_rate,
                                 color = CreditRating)) +
  ggtitle("Avg. Borrower Rate by Term & Credit Rating") +
  geom_line(aes(group = CreditRating), size = 1.0) +
  scale_color_brewer(palette = "Set1", direction = -1) +
  scale_y_continuous(breaks = seq(0, 0.35, 0.05)) +
  geom_point() +
  theme(legend.position="bottom", axis.text.x = element_text(angle = 90, hjust = 1)) +
  guides(colour = guide_legend(nrow = 1)) +
  facet_wrap(~Term, ncol = 1)

```

* Ok looks like, 12-month and 60-month loans have only been given out since Q4
2010 / Q1 2011. (Interestingly starting with borrowers with a good creddit 
rating, before they were accessible for borrowers of all credit classes).
* What is striking about the plot is the jump in Q2 2009 for the rate of
borrowers with a D credit rating. The rates of all other credit classes slightly
decrease in that quarter. A reason for this odd feature might be the change
between the two types of credit scores which happened during that time and 
which we combined in our data.
* Lets focus in the next plot on the time from 2011 onwards, during which all 
3 types of loans were offered.

```{r echo=FALSE, fig.width = 6, fig.height=6}
ggplot(subset(rate_by_CreditRating, LoanOriginationQuarter > "2010 Q4"), 
       aes(x = LoanOriginationQuarter,
                                 y = avg_rate,
                                 color = CreditRating)) +
  ggtitle("Avg. Borrower Rate by Term & Credit Rating (2011-2014)") +
  geom_line(aes(group = CreditRating), size = 1.0) +
  scale_color_brewer(palette = "Set1", direction = -1 ) +
  scale_y_continuous(breaks = seq(0, 0.35, 0.05)) +
  geom_point() +
  theme(legend.position="bottom", axis.text.x = element_text(angle = 90, hjust = 1)) +
  guides(colour = guide_legend(nrow = 1)) +
  facet_wrap(~Term, ncol = 1)
```

Adding a summary of the average rates per loan (data from 2011 onwards):

```{r echo=FALSE, Summary_rates} 
subset(prosper, LoanOriginationQuarter > "2010 Q4") %>%
  group_by(Term, CreditRating) %>%
  summarise(avg_rate = mean(BorrowerRate))

```

## What factors influence the borrower rate?

First, lets see how the employment status affects the borrower rate:

```{r echo = FALSE, Employment}
# Setting empty EmploymentStatus to NA
prosper$EmploymentStatus[prosper$EmploymentStatus == ""] <- NA 

#Converting to a data.table, for easier subsetting
ggplot(subset(prosper, !is.na(EmploymentStatus)), 
       aes(x = EmploymentStatus, y = BorrowerRate)) +
  geom_boxplot()
```

* Although there is some overlap in the different definitons  of EmploymentStatus,
it becomes clear that the median borrower rate for unemployed borrowers is
higher.

Does it affect the BorrowerRate if the borrower owns a house?

```{r echo=F, HomeOwner}
ggplot(prosper, 
       aes(x = IsBorrowerHomeowner, y = BorrowerRate)) +
  geom_boxplot()
```

* HomeOwners get a lower rate then people not owning a house.
This makes sense, because homeowners can always sell or put a mortage on their 
homes.

Next, let's take the view of the lenders. What return can they excpect?

```{r echo=F, Return}
ggplot(prosper, aes(x = LenderYield)) +
  geom_histogram(binwidth = 0.01) + 
  scale_x_continuous(lim = c(0, 0.400))

summary(prosper$LenderYield)
```

* The lender yield is the interest rate minus the expected fee payments. 
* It is one of the most important metrics for lenders.
* As we can see, the median lender yield is ~17%. The distribution seems to be
bimodal, peaking again at 31%. We will look further into this at a later point.



## Correlations between different features

Because, the dataset contains over 80 different variables, let's first subset it
to the ones most relevant to our investigation:

```{r}
prosper_sub <- prosper[c("BorrowerRate", "CreditRating","ProsperScore", "Term", "BorrowerAPR",
                         "DebtToIncomeRatio", "StatedMonthlyIncome", "AvailableBankcardCredit", "EmploymentStatus")]
```

Next, let's use the `GGally` package to create a correlation matrix of these
variables. This wil give us a good overview:

```{r echo = FALSE}
ggpairs(prosper_sub)
```



```{r echo=FALSE, ProsperScore}
ggplot(prosper, aes(x = ProsperScore, 
                     y = BorrowerRate)) +
  ggtitle("Borrower rate by ProsperScore") +
   geom_jitter() +
  geom_smooth(method = "lm")
```

```{r}
cor(prosper$BorrowerRate, prosper$ProsperScore, use = "complete.obs")
```

* As the plot and the calculation of the correlation coefficient show, the Prosper
Score and the Borrower Rate are negatively correlated (-0.65). This makes sense,
since the Prosper Score is baded on the Credit Rating of the individual. With a 
better (higher) score, the risk is lower and a better rate can be offered.


## Multivariate Plots

Let's further investigate the relationship between a borrowers prosper score, 
credit rating and borrowing rate:

```{r echo=FALSE, RateScoreRating}
ggplot(prosper, aes(x = ProsperScore, 
                     y = BorrowerRate,
                    color = CreditRating)) +
  ggtitle("Borrower rate by ProsperScore and Credit Rating") +
  geom_jitter(alpha = 0.5) +
  scale_x_continuous(breaks = seq(1,11,1)) +
  scale_y_continuous(breaks = seq(0.0, 0.4,0.05), limits = c(0,0.4)) +
  scale_color_brewer(palette = "Set1", direction = -1) +
  geom_smooth(method = "lm", color = "black")
```

* Again, we can see that persons with an higher Prosper Score revieve the better
borrowing rates.
* By adding a coloring by Credit Rating, we can see a clear differentiation in 
the borrowing rates of the different credit ratings. 
* What is interesting is that prosper seems to set certain max. borrowing rates
for the different credit classes. Therefore in the plot there is a clear cut at 
a borrowing rate 0f 36%. Only a view outliers have a borrowing rate above that 
level.

Next, let's come back to Prospers overall performance over the years. We have
seen before, that the total amount of loans originated has increased over the 
years. But what is even more relevant is the development of defaulted vs. currrent
loans. 

```{r echo=FALSE, fig.width = 12, OriginatedVsDelinquent}
originatedVsDelinquent_by_quarter <- subset(prosper, 
                                            LoanOriginationQuarter > "2008 Q1") %>%
  group_by(LoanOriginationQuarter) %>%
  summarise(totalOriginated = sum(LoanOriginalAmount),
            totalDelinquent = sum(AmountDelinquent))

# change structure of dataframe to prepare for paired bar plot
originatedVsDelinquent_by_quarter_MELT <- melt(originatedVsDelinquent_by_quarter, id.vars= "LoanOriginationQuarter")

ggplot(originatedVsDelinquent_by_quarter_MELT, aes(x = LoanOriginationQuarter, y = value, fill = variable)) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_fill_manual(labels = c("originated", "delinquent"), values = c("darkgreen", "darkred")) +
  scale_y_continuous(
    breaks = seq(0, 150000000, 25000000),
    labels = dollar_format(scale = 0.000001, suffix = "M")
  ) +
  labs(
    title = "Originated vs. Delinquent Loan Amount per Quarter",
    x = "Quarter", y = "Total amount"
  ) +
  theme(legend.title = element_blank(), legend.position = "bottom", axis.text.x = element_text(angle = 90, hjust = 1)) +
  guides(colour = guide_legend(nrow = 1))
```

## Building a linear model to predict the Borrower Rate
```{r echo=FALSE}
library(memisc)
```

```{r}
# Building the model, bottom up approach
m1 <- lm(BorrowerRate ~ CreditRating, data = prosper)
m2 <- update(m1, ~ . + StatedMonthlyIncome)
m3 <- update(m2, ~ . + IsBorrowerHomeowner)
m4 <- update(m3, ~ . + EmploymentStatus)
m5 <- update(m4, ~ . + AvailableBankcardCredit)
m6 <- update(m5, ~ . + DebtToIncomeRatio)

# Summary of the r-squared tests of the different models  
mtable(m1,m2, m3, m4, m5, m6)
```

The model takes 6 variables into account: 
* The credit rating (the variable we created earlier),
* the monthly income,
* if the persons own a house or not,
* the employment status,
* the available bankcard credit,
* and the persons debt to income ratio.
* Taking this 6 variables into acount we get a R-squared value of 0.816. This
is pretty good. (A r-square value of 1 would mean that our fitted regression line
would explain 100% of the datapoints.)

Lets test the model using a sample loan from our data:
```{r}
set.seed(7262) # Setting a seed for reproduable result, remove line to use general model

thisloan = sample(prosper, 1)

thisloan[c(
  "BorrowerRate", "CreditRating", "StatedMonthlyIncome",
  "IsBorrowerHomeowner",
  "EmploymentStatus", "AvailableBankcardCredit", "DebtToIncomeRatio"
)]

modelEstimate = predict(m6, newdata = thisloan, interval = "prediction",
                        level = .95)
modelEstimate

```

* Here, we can see that the sample loan randomly selected from the dataset has
a bowwowing rate of `0.1419`. Our model predicted a borrowing
rate of `0.1540776`. This is slightly higher, but well withing the computed ranges
of the 95% confidence interval.


## Final Plots

### 1. Originated vs. Delinquent Loans by Quarter
```{r echo=FALSE, Final1}
ggplot(originatedVsDelinquent_by_quarter_MELT, aes(x=LoanOriginationQuarter, y=value, fill=variable)) +
    geom_bar(stat="identity", position="dodge") +
    scale_fill_manual(labels = c("originated", "delinquent"), values = c("darkgreen", "darkred")) +
    scale_y_continuous(
      breaks = seq(0, 150000000, 25000000), 
      labels = dollar_format(scale = 0.000001, suffix = "M")
      ) +
  labs(title = "Originated vs. Delinquent Loan Amount per Quarter",
       x = "Quarter", y = "Total amount") +
   theme(legend.title=element_blank(), legend.position="bottom", axis.text.x = element_text(angle = 90, hjust = 1)) +
  guides(colour = guide_legend(nrow = 1)) + 
  annotate("segment", x = "2009 Q2", xend = "2009 Q2", y = 0, yend = 162000000,
  colour = "grey",size = 1.5) +
  annotate("text",  label = "Relaunch", x = "2009 Q4", y = 160000000, colour = "grey")

```

* This plot really captures the development of Prosper. It shows how the site managed
to grow the amount of loans orgininated year by year, while maintaining a low level of
loans beeing delinquent. After a short decrease prior to the relaunch of the site in 
2009 there has been a steady upwards trend.

### 2. Loan Issuance By State

```{r echo=FALSE, Final2}
ggplot(state_data, aes(fill = total_loans_issued)) +
  ggtitle("Loan Issuance By State") +
  geom_map(aes(map_id = abbr2state(BorrowerState)), map = map) +
  expand_limits(x = map$long, y = map$lat) +
  scale_fill_distiller(
    direction = 1,
    guide = "legend",
    name = "Total Loans",
    labels = dollar_format(scale = 0.000001, suffix = "M")) +
  theme_void() +
  theme(plot.title = element_text(hjust = 0.5)) 
```

* This plot captures in which states prosper is particulary active. California
is thei central hub of business region. One reason for this might be that the
company is baded in this area.

### 3. Borrower rate by ProsperScore and Credit Rating

```{r echo=FALSE, Final3}
ggplot(prosper, aes(x = ProsperScore, 
                     y = BorrowerRate,
                    color = CreditRating)) +
  ggtitle("Borrower rate by ProsperScore and Credit Rating") +
  geom_jitter(alpha = 0.5) +
  scale_x_continuous(breaks = seq(1,11,1)) +
  scale_y_continuous(breaks = seq(0.0, 0.4,0.05), limits = c(0,0.4)) +
  scale_color_brewer(palette = "Set1", direction = -1) +
  geom_smooth(method = "lm", color = "black")
```

* This graph clearly highlights the relationsship between a persons credit rating,
prosper score and the borrower rate they recieved. 
* Persons with a better Credit Rating and hence ahigher Prosper Score recieve 
the better borrowing rates.
* The straight horizontal lines in the graph indicate that there are maximum
rates for the different credit ratings set by Prosper.

## Reflection

By exploring and analyzing the dataset of Propser, I have been able to gain 
insights about the P2P lending business in general and also the business
development of Prosper.com. One of the main metrics of interest for loan applicants
and lenders is the borrowing rate. I was able to uncover the key factors that
influence the interest rate an individual recieves. Furthermore I was able
to build a linear model to predict the borrowing rate one recieves,
taking into account the credit rate, income, employment status, 
home-ownership and credit and debt info of an individual. 

One of the main challenges was to get familiar with the different variables in
the data. With 81 variables per record, exploratory data analysis and plots are
particulary important to gain an understanding of the data and to uncover
relationships between different variables. R has proven to be a very effective.
here. 

A promising avenue of future research on the dataset is the analysis of 
factors which are signs that a loan will not be payed back fully or on time. 
Building a model which could predict the likelyhood of delinquent loans would
greatly reduce the risk taken by the lender and would be a valuable tool for
businesses such as Prosper.com.

